---
title: "Code used to data wrangle and create visualizations on a popular games data set"
author: "Faye, Genevieve, Matt"
date: '2023-05-18'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Into the GameR-verse

Welcome! In this notebook our team dives into the GameR-verse by using R code to analyze a data set on the most popular video games from 1980 to 2023.

This publicly available data set can be accessed from the Kaggle website using the link below: https://www.kaggle.com/datasets/arnabchaki/popular-video-games-1980-2023

## Setting Up the Data

Use getwd() to make sure you're in the same directory as games.csv. If not, use getwd("/...") to get into the same directory.

Installing and loading in stringr and dplyr for cleaning data.

```{r}
# Uncomment the install commands below if these libraries have never been installed yet. If they are already installed, you must keep these commented otherwise you'll run into errors
#install.packages("dplyr") 
#install.packages("stringr")

library(dplyr)
library(stringr)
library(tidyr)
library(data.table) 
```

## Reading in the data

Use the read.csv() function to read in the csv as a dataframe. Make sure games.csv is within the same folder as this RMD file. If it's not, you will need to specify a specific path to the csv file.

```{r}
games_df <- read.csv("games.csv")
```

## Cleaning up the data

Changing all values that have a "K" into numbers by using stringr's str_replace() function. str_replace() takes the column, the string values that need to be replaced, and what to replace that string value with.\
\
In this line, we're targeting all the "K"s in the Times Listed column, since we want 3900 as a numeric value and not "3.9K". The "\$" is used to mark the end of the string and doesn't confuse R. Afterwards, as.numeric() is used to change the "3.9" into 3.9 (a number and not a string) and multiplied by 1000.

```{r}
games_df$Times.Listed <- as.numeric(str_replace(games_df$Times.Listed, "K$", "")) * 1000
```

Doing the above for columns with "K" values (Number of Reviews, Plays, Playing, Backlogs, and Wishlist)

```{r}
games_df$Number.of.Reviews <- as.numeric(str_replace(games_df$Number.of.Reviews, "K$", "")) * 1000
games_df$Plays <- as.numeric(str_replace(games_df$Plays, "K$", "")) * 1000
games_df$Playing <- as.numeric(str_replace(games_df$Playing, "K$", "")) * 1000
games_df$Backlogs <- as.numeric(str_replace(games_df$Backlogs, "K$", "")) * 1000
games_df$Wishlist <- as.numeric(str_replace(games_df$Wishlist, "K$", "")) * 1000
```

Make sure ***not to run these chunks twice*** -- otherwise you'll multiply the values by 1000000. If that happens type rm(games_df) into the console and rerun the chunks.

## Creating new "Month" and "Year" columns

Installing and loading in lubridate for formatting.

```{r}
#install.packages("lubridate")
library(lubridate)
```

Converting Date column into a date format. Originally, the dates were in this format: "Feb 25, 2022". After using mdy() to change into 2022-02-25 so that we can use month() and year().

-   Warning: 3 failed to parse -- we will ignore these 3 for now

```{r}
games_df$Release.Date <- mdy(games_df$Release.Date)
```

Using month() to create a month column and year() to create a year column.

```{r}
games_df$Release.Month <- month(games_df$Release.Date, label = TRUE) 
#label = TRUE is used to get month name rather than the number (February vs 02) -- we can get 02 if we delete label

games_df$Release.Year <- year(games_df$Release.Date)
```

## Plots

Installing and loading in ggplot for visualization.

theme(axis.text.x = element_text(angle = 45, hjust = 1),

panel.grid = element_blank())

```{r}
#install.packages("ggplot2")
library(ggplot2)
```

Right now, the Genres column is formatted in vectors, which hold multiple genres in one row. We need to separate each genre into its own separate row to make it easier for ggplot2 to plot each genre.

```{r}
# convert genre column into string
games_genres <- games_df # make a copy of the original df
games_genres$Genres <- as.character(games_genres$Genres) 
games_genres[games_genres==""]<-NA # replace blank cells with N/A

# for loop that separates each genre into its own row
for (row in 1:nrow(games_genres)) { # do this for all rows in the df
  new_str <- gsub("\\[|\\]", "", games_genres$Genres[row]) # remove the brackets 
  new_str <- gsub("\\'|\\'", "", new_str) # remove the "" 
  new_str <- gsub(" ", "", new_str) # remove any extra spaces
  games_genres$Genres[row] <- new_str # the cleaned up string will be replaced 
  # old genre format = "['Genre1', 'Genre2']" --> new genre format = "Genre1, Genre2"
}

# split into multiple rows using the "," 
games_genres = separate_rows(games_genres,"Genres",sep = ",")
games_genres <- games_genres %>% drop_na(Rating)
```

Plotting Most Popular Genres by Year

```{r}
# make a df which we'll use to plot genre vs years
genre_year_df <- games_genres %>% 
  group_by(Genres, Release.Year)

genre_year_counts <- genre_year_df %>% 
  summarize(Count = n())

most_popular_by_year <- genre_year_counts %>% 
  group_by(Genres) %>% 
  top_n(1, Count) %>% 
  ungroup()

# remove the top three rows which were blank 
most_popular_by_year <- most_popular_by_year[-c(1,2,3),] 
```

```{r}
ggplot(most_popular_by_year, aes(x = Release.Year, y = Count, fill = Genres)) +
  geom_col() +
  labs(title = "Most Popular Genres by Year", x = "Year", y = "") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Plotting Average Rating per Genre

```{r}
# make a df which is grouped by Genre and has the average rating per genre
genres_rating_avgs <- games_genres %>%
  group_by(Genres) %>%
  summarise_at(vars(Rating), list(avg_rating = mean)) 

# remove the top row which was blank
genres_rating_avgs <- genres_rating_avgs[-c(1),] 
```

```{r}
ggplot(data=genres_rating_avgs, aes(x = reorder(Genres, -avg_rating), y=avg_rating)) +
  geom_bar(stat="identity", width=0.5, fill="plum4") +
  theme(axis.text.x = element_text(angle = 65, hjust = 1), panel.grid = element_blank()) +
  labs(title = "Average Rating per Genre",
  x = "Genre", y = "Average Rating (out of 5)") 
```

Plotting Average Plays per Genre

```{r}
# make a df which is grouped by Genre and has the average plays per genre
genres_plays_avgs <- games_genres %>%
  group_by(Genres) %>%
  summarise_at(vars(Plays), list(avg_plays = mean)) %>% 
  drop_na(avg_plays)

# remove the top row which was blank
genres_plays_avgs <- genres_plays_avgs[-c(1),]

# try to find the blank row
games_genres[games_genres$Genres %like% " ", ]
print(genres_plays_avgs[1,1])
```

```{r}
ggplot(data=genres_plays_avgs, aes(x = reorder(Genres, -avg_plays), y=avg_plays)) +
  geom_bar(stat="identity", width=0.5, fill="plum4") +
  theme(axis.text.x = element_text(angle = 65, hjust = 1), panel.grid = element_blank()) +
  labs(title = "Average Plays per Genre",
  x = "Genre", y = "Average Plays") 
```

## 

```{r}
```

## 

```{r}
```

## 

```{r}
```

## 

```{r}
```

## 

```{r}
```

## 

```{r}
```

## 

```{r}
```

## 

```{r}
```

## 

```{r}
```

## 

```{r}
```
